---
title: "Lasso Analysis To Find Regulatory K-mers in MZT"
author: "Santiago Medina"
date: "3/22/2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )
library(tidyverse)
library(ggthemes)
```

## Overview

I run a lasso analysis to find regulatory k-mer with a response variable and a set of predictors.

Response: mRNA stability indicator (decay rate), this indicator was estimated with a linear model, TPM  ~ time.
Predictors: 2 PLSs optimality components and all the possible 6-mers.

This analysis is run 4 times:

+ all maternal genes (~4k) for the Ribo0 and PolyA data sets.
+ all maternal genes and no MiR-430 6-mer sits (~3k) for he Ribo0 and PolyA data sets.

In each of the cases above, we performed 6-Kfold validation to obtain reliable estimates.

The output of this analysis if a regularization path (with the LASSO) and the cross-validation for each fold to detect the best $\lambda$ parameter that maximizes the R2 score.


## Lasso Paths

```{r lassopaths, fig.height=6, fig.width=8}
lasso_paths <- read_csv("results_data/lasso_paths.csv")

lasso_paths <- lasso_paths %>% 
  rename(vars = X1) %>% 
  gather(alpha, coef,-vars, -kfold, -sample_condition, -which_genes) %>% 
  mutate(alpha = as.numeric(alpha))

coef_type <- function(vname) {
  if (vname == "PLS1") return("PLS1")
  if (vname == "PLS2") return("PLS2")
  if (vname == "GCACTT") return("MiR-430")
  else return("k-mer")
}

lasso_paths <- lasso_paths %>% 
  mutate(vartype = map_chr(vars, coef_type))

lasso_paths %>% 
  filter(kfold == 1) %>% 
  ggplot(aes(x=alpha, y=coef, color=vartype, group=vars)) +
  geom_line(aes(size=vartype)) +
  scale_x_log10() +
  geom_hline(yintercept = 0, linetype=3) +
  scale_size_manual(values = c(1/25, 1, 1, 1/2)) +
  scale_color_colorblind() +
  facet_grid(sample_condition ~ which_genes)

```

## Best $\lambda$

```{r best_lambda, fig.height=6, fig.width=7}
cval <- read_csv("results_data/lasso_cross_val_params.csv")
cval %>% 
  filter(mean_test_score > -.10) %>% 
  ggplot(aes(x=param_alpha, y=mean_test_score * 100, color=which_genes)) +
  geom_point() +
  geom_smooth() +
  scale_x_log10() +
  facet_grid(~sample_condition)
```

For each $\lambda$ in the plot above we have six points coming from the 6s fold in CV.
Now let's get the best lambda value for each case.


```{r}
# get the best lambda value by taking the median across cases
best_lambda <- 
  cval %>% 
  group_by(
    param_alpha, which_genes, sample_condition
  ) %>% 
  summarise(
    mean_test_score = median(mean_test_score)
  ) %>% 
  ungroup() %>% 
  group_by(sample_condition, which_genes) %>% 
  nest() %>% 
  mutate(best_l = map(data, function(x) arrange(x, -mean_test_score) %>% slice(1:1))) %>% 
  unnest(best_l) %>% 
  select(-data)

knitr::kable(best_lambda)
```

```{r}
# get the coefs for thos lambdas
results <- best_lambda %>% 
  inner_join(lasso_paths, by = c("sample_condition", "which_genes")) %>% 
  filter(alpha == param_alpha)

# check elements that pass the filter -------------------------------------
# the filter is that for each case, in the the fold the coefficient should
# not be zero

candidates <- 
  results %>% 
  group_by(sample_condition, which_genes, vars) %>% 
  nest() %>% 
  mutate(
    passesfilter = map_lgl(data, function(df) all((abs(df$coef) > 0)))
  ) %>% 
  filter(passesfilter)

# now get the median for each variable
candidates <- 
  candidates %>% 
  select(-passesfilter) %>% 
  unnest() %>% 
  ungroup() %>% 
  group_by(sample_condition, which_genes, vars) %>% 
  summarise(effect = median(coef)) %>% 
  ungroup()

```

```{r candidates, fig.height=7, fig.width=3}
# rank base on mean effect across conditions ------------------------------
ranking <- 
  candidates %>% 
  group_by(vars) %>% 
  summarise(me = median(effect)) %>% 
  arrange(me) %>% 
  pull(vars)

candidates$vars <- factor(candidates$vars, levels = ranking)
  
candidates %>% 
  mutate(
    ef = sign(effect),
    ef = map_chr(ef, ~if_else(. == 1, "stabalizing", "destabalizing")),
    ef = factor(ef, levels = c("stabalizing", "destabalizing"))
  ) %>% 
  ggplot(aes(x=which_genes, y=vars, fill=effect)) +
  geom_tile() +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  facet_grid(ef~sample_condition, scales = "free_y", space = "free_y") +
  scale_fill_viridis_c() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(
    y = NULL,
    x = NULL,
    title = "Candidates LASO analysis"
  )
```

**NOTE**: If a k-mer is not significant for the condition in the heatmap above there is not color.

```{r}
# save results
write_csv(candidates, "results_data/candidate_elements_laso_analysis.csv")
```

