---
title: "Evaluatio the Model in WT (MZT) conditions"
author: "Santiago Medina"
date: "5/5/2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )

library(tidyverse)
library(ggthemes)
library(caret)
theme_set(theme_tufte(base_family = "Helvetica"))
```

This analysis is done for mainly maternal genes.

## The data sets
1. Model Predictions
2. log2FC data
3. Optimality (PLS1 or categories from Ariel's embo2016)
4. MiR-430 sites

## 1. Model Predictions Bassed on the Codon Compositio

```{r}
genes_to_predict <- bind_rows(
  read_csv("../19-01-22-PredictiveModelMrnaStability/results_data/test_data.csv"),
  read_csv("../19-01-22-PredictiveModelMrnaStability/results_data/train_data.csv")
)

mdl <- read_rds("../19-01-22-PredictiveModelMrnaStability/results_data/trained_models/svmRModel.rds")

genes_to_predict$predictedDecay <- predict(mdl, newdata = genes_to_predict)

# subset the important columns
stabCodons <- 
  genes_to_predict %>% 
  select(Gene_ID, predictedDecay)
```


## 2. Log Fold Change Data
```{r}
log2data <- read_csv("../../data/19-02-05-FoldChangeData/data/log2FC_earlyVSlate_tidytimecourse.csv")
```

## 3. Optimality AND MIR-430 data

```{r}
programs <- read_csv("../19-02-24-OverlapPathwaysFig3/results_data/regulatory_pathways_matrix.csv") %>% 
  filter(is_maternal == TRUE) %>% 
  select(Gene_ID, m6A, MiR430, PLS1, PLS2)


## define the top 500 optimal and non-optimal genes

top_optimal <- 
  programs %>% 
  filter(MiR430 == 0) %>% 
  arrange(-PLS1) %>%
  pull(Gene_ID) %>% 
  .[1:500]

top_nonoptimal <- 
  programs %>% 
  filter(MiR430 == 0) %>% 
  arrange(PLS1) %>%
  pull(Gene_ID) %>% 
  .[1:500]

mir430 <- programs %>% filter(MiR430 > 0) %>% pull(Gene_ID)
optimalityclass <- bind_rows(
  tibble(Gene_ID = top_optimal, class = "optimal"),
  tibble(Gene_ID = top_nonoptimal, class = "non-optimal"),
  tibble(Gene_ID = mir430, class = "MiR-430")
)

programs <- left_join(programs, optimalityclass) %>% 
  replace_na(list(class="neutral"))
```
## Fig 2

```{r fig02, fig.height=2.5, fig.width=7}
data <- inner_join(programs, stabCodons) %>% 
  inner_join(log2data)


# show the median values
medians <- data %>% 
  filter(time == 6) %>% 
  group_by(sample_condition, class) %>% 
  summarise(m_predictedDecay = mean(predictedDecay), m_log2FC = mean(log2FC))


data %>% 
  filter(time == 6) %>% 
  ggplot(aes(x=predictedDecay, y=log2FC, color=class)) +
  geom_point(alpha=3/5, shape=16, size=1) +
  geom_point(data = medians, aes(x=m_predictedDecay, y=m_log2FC), size=3.5, shape=16, color="black", alpha=.95) +
  geom_point(data = medians, aes(x=m_predictedDecay, y=m_log2FC), size=3, shape=16, alpha=.95) +
  ggpubr::stat_cor(color="black", size=2.5) +
  geom_rangeframe(color="black") +
  scale_color_manual(values = c(NA, "grey", "blue", "red")) +
  facet_grid(~sample_condition) +
  labs(
    title = "all maternal genes ~5K (miR-430 not shown)",
    x = "predicted mRNA stability bassed on codon composition",
    y = "log 2 fol change WT\n(6hrs / 3hrs)"
  )
```

```{r fig02aAllGenes, fig.height=2.5, fig.width=7}
data %>% 
  filter(time == 6) %>% 
  ggplot(aes(x=predictedDecay, y=log2FC, color=class)) +
  geom_point(alpha=3/5, shape=16, size=1) +
  geom_point(data = medians, aes(x=m_predictedDecay, y=m_log2FC), size=3.5, shape=16, color="black", alpha=.95) +
  geom_point(data = medians, aes(x=m_predictedDecay, y=m_log2FC), size=3, shape=16, alpha=.95) +
  ggpubr::stat_cor(color="black", size=2.5) +
  geom_rangeframe(color="black") +
  scale_color_manual(values = c("forestgreen", "grey", "blue", "red")) +
  facet_grid(~sample_condition) +
  labs(
    title = "all maternal genes ~5K",
    x = "predicted mRNA stability bassed on codon composition",
    y = "log 2 fol change WT\n(6hrs / 3hrs)"
  )
```

## Residual Plots To show MiR-430 are over-estimated

To get the residual I will fit a linear model:

$$
log2fc\text{(Xhrs / 3hrs)} = \alpha + \beta\;\hat{y}
$$
Where $\hat{y}$ is the predicted stability bassed on the codon composition. The linear model is to 
bassically put the predictions in the same scale as the log2FC.


```{r residuals_M6A, fig.height=5, fig.width=3.5}
preds <- 
  data %>% 
  filter(time > 4) %>% 
  filter(!is.infinite(log2FC)) %>% 
  group_by(sample_condition, time) %>% 
  nest() %>% 
  mutate(
    fit = map(data, ~lm(log2FC ~ predictedDecay, data = .)),
    preds = map(fit, broom::augment)
  ) %>% 
  unnest(preds) %>% 
  inner_join(data, by=c("time", "sample_condition", "log2FC", "predictedDecay"))


preds %>% 
  filter(time %% 1 == 0, time < 8) %>% 
  mutate(time = paste0(time, " hrs")) %>% 
  ggplot(aes(x=m6A, y=.resid, fill = m6A)) +
  geom_boxplot(outlier.shape = NA, size = 1/3) +
  geom_rangeframe(alpha=1/3) +
  facet_grid(time~sample_condition) +
  scale_fill_manual(values = c("grey", "steelblue")) +
  scale_x_discrete(labels=c("no targets", "m6A targets")) +
  scale_y_continuous(breaks = c(-3, 0, 3)) +
  coord_cartesian(ylim = c(-6, 4)) +
  labs(
    x = NULL,
    y = "residual\n(observed - predicted)",
    subtitle = "Model Overestimates m6A targets in WT"
  ) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 60, hjust = 1))
```


```{r residuals_MiR430, fig.height=5, fig.width=4}
preds %>% 
  filter(time %% 1 == 0, time < 8, MiR430 < 5) %>% 
  mutate(time = paste0(time, " hrs")) %>% 
  ggplot(aes(x=MiR430, y=.resid, group=MiR430, fill = as.character(MiR430))) +
  geom_boxplot(outlier.shape = NA, size = 1/3) +
  geom_rangeframe(alpha=1/3) +
  facet_grid(time~sample_condition) +
  scale_y_continuous(breaks = c(-3, 0, 3)) +
  scale_fill_manual(values = c("grey", "#074050", "#105965", "#217a79", "#4c9b82")) +
  coord_cartesian(ylim = c(-6, 4)) +
  labs(
    x = "# sites with MiR-430 seed",
    y = "residual\n(observed - predicted)",
    subtitle = "Model Overestimates MiR-430 targets in WT"
  ) +
  theme(legend.position = "none")
```

