---
title: "Codon composition strongly predicts mRNA stability"
author: "Santiago Medina"
date: "March 1, 2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/f01_", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )
library(tidyverse)
library(ggthemes)
library(caret)

theme_set(theme_tufte(base_family = "Helvetica"))
```

##	Half-life or decay distribution (Santi)

```{r decaydistribution, fig.height=2, fig.width=3}
# the beta parameter is the decay rate

decay_aa <- 
  read_csv("../../../../results/19-01-11-GetDecayRateFromTimeCourse/results_data/estimated_decay_rates.csv") %>% 
  filter(term == "beta")

decay_aa %>% 
  ggplot(aes(x=estimate)) +
  geom_histogram(bins = 100) +
  labs(
    x = "decay rate",
    title = "distribution of decay in alpha-Amanitin "
  ) +
  geom_rangeframe() +
  theme(text = element_text(size = 8))

```

## C)	Model to predict mRNA decay in alpha (scatter plot, R and P value), How much can you explain (Santi)


Best tuning parameters
```{r tuning}
mdl <- read_rds("../../../../results/19-01-22-PredictiveModelMrnaStability/results_data/trained_models/svmRModel.rds")
mdl$bestTune
```

```{r mdl_test_data_predictions, fig.height=3, fig.width=3}
# get a confidence interval
r2_observed <- mean(mdl$resample$Rsquared)
ci_r2 <- broom::tidy(t.test(mdl$resample$Rsquared), confint = T)

## test data
test_data <- read_csv("../../../../results/19-01-22-PredictiveModelMrnaStability/results_data/test_data.csv")

test_data$prediction <- predict(mdl, test_data)


test_data %>% 
  ggplot(aes(prediction, decay_rate)) +
  geom_point(size = 1/2, color = "steelblue") +
  geom_abline(linetype = 3) +
  ggpubr::stat_cor(size = 2) +
  geom_rangeframe() +
  labs(
    subtitle = str_c("R2 = ", round(r2_observed, 3), " (", round(ci_r2$conf.low, 3), ", ", round(ci_r2$conf.high, 3), ") 95% CI"),
    title = "Predictive Model Codon Composition",
    x = "predicted",
    y = "observed"
  ) +
  theme(text = element_text(size = 8))
```


D)	Calculate the mRNA decay of the 1nt out of frame (codon vs nt) and Gopal can inject those. (Santi and Gopal)

TODO: tibble with reporter sequences
I do not think that the reporter sequences I am using from the EMBO paper are correct.


```{r reporters_predictions, fig.height=3, fig.width=2}
reporters <- readxl::read_excel("results_data/reporters_sequences.xlsx")

reporter_cc <- 
  reporters %>% 
  mutate(
    cc = map(sequence, codonr::count_codons)
  ) %>% 
  select(reporter_id, cc) %>% 
  rename(Gene_ID = reporter_id) %>% 
  unnest(cc) %>% 
  spread(key = codon, value = n) %>% 
  replace(., is.na(.), 0)

reporter_cc$prediction <- predict(mdl, newdata = reporter_cc)


xsub <- reporter_cc %>% 
  select(Gene_ID, prediction) %>% 
  filter(str_detect(Gene_ID, "cherr"))

mdl$trainingData %>% 
  mutate(prediction = predict(mdl, newdata = .)) %>% 
  ggplot(aes(x="", y = prediction)) +
  geom_point(position = "jitter", color = "grey90", shape = ".") +
  geom_point(data = xsub, aes(x="", y=prediction), color = "steelblue") +
  ggrepel::geom_text_repel(data = xsub, aes(x="", y=prediction, label = Gene_ID), size = 2) +
  labs(
    x = NULL,
    y = "predicted mRNA stability"
  )
```
