---
title: "Workflow to process arielome for analysis"
author: "Santiago Medina"
date: "6/8/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
dirtoraw <- "../../../data/19-03-23-Arielome/processed_arielome_data/"
meta <- read_csv("../../../data/19-03-23-Arielome/Arielome/metadata.csv") %>% 
  filter(condition == "wt")

```

The arielome data is located at: *../../../data/19-03-23-Arielome/processed_arielome_data/*. The files in that directory with termination *\*premature_stop.csv* are the sequences from the library containing a premature stop codon and possibly a 3' UTR sequence. Plese see the *../../../data/19-03-23-Arielome/README.md* for further reference.



The snakemakw workflow has two steps:

1. `filter_by_seqlen`: Remove short seqs (check the file to check params).
2. `add_optimality`: Use the predictive model to predict the stability of the sequences. I will use the prediction as an indicator of codon optimality. Here, I also add an indicator variable for the presence of miR-430 6-mer.

I will only process the files at 2hrs and 8hrs without treatment.

```{r}
knitr::kable(meta)
```

To run this pipeline use:

```bash
snakemake --cores 8 data/arielome_opt_mir.csv
```

The main output of this pipeline is the file `data/arielome_opt_mir.csv`.

```{r echo=FALSE, message=FALSE}
mout <- read_csv("data/arielome_opt_mir.csv")
sample_n(mout, 10) %>% 
  knitr::kable()
```

