---
title: "EDA, Positive Controls"
author: "Santiago Medina"
date: "6/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )

library(tidyverse)
library(ggridges)
library(gam)

d <- read_csv("../200608-process-arielome/data/arielome_opt_mir.csv")

tmp_mo <- d %>% 
  filter(species == "fish", condition != "wt")

tmp_wt <- d %>% 
  filter(species == "fish", time == 2)

d <- bind_rows(tmp_mo, tmp_wt)
```

## Overview

Some note from an email I send to Ariel:

1. In the sequences that I have filter. Can I see the regulation of codon optimality (The variable I am using as an indicator of optimality)?
2. Can I see miR-430 regulation in these sequences (6mer, 7mer, and 8mer)? 
3. What is the complexity of these sequences? If I see a lot of repeats (no complexity) probably I won't see the effect.

If the data passes these controls I will be able to say something.

## EDA

```{r}
d %>% 
  count(seq_id) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 5) +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(from=0, to=1500, by=250)) +
  labs(
    x = "sequences is represented n-times"
  )
```

We have some sequences that are repeated muiltiple times. I will discard sequences that are repeated more than 30 times.

```{r}
repeated_more_than_30 <- 
  d %>% 
  count(seq_id) %>% 
  filter(n > 20) %>% 
  pull(seq_id)

d <- filter(d, !seq_id %in% repeated_more_than_30)

d %>% 
  count(seq_id) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1) +
  scale_y_log10() +
  scale_x_continuous() +
  labs(
    x = "sequences is represented n-times",
    y = "repeats removed"
  )
```

### Can we see codon optimality regulation?

First I plot the destribution of optimality scores.

```{r}
d %>% 
  ggplot(aes(x = optimality, fill = as.character(time))) +
  geom_histogram(bins = 100) +
  facet_wrap(~sample_name) +
  labs(
    color = "Time [hrs]",
    x = "Predicted stability (codon optimality)"
    )
```

Observation: Maybe we don't find sequences that are very optimal/non-optimal.
Why?

$$
\text{Codon composition arielome } != \text{Codon composition endogenous}
$$

```{r}
tmp <- d %>% 
  mutate(time = paste0(time, " hrs")) %>% 
  group_by(time) %>% 
  summarise(opt = median(optimality))

d %>% 
  mutate(time = paste0(time, " hrs")) %>% 
  ggplot(aes(x= optimality, fill = time)) +
  geom_density(alpha = .5, color = "white") +
  geom_vline(data = tmp, aes(xintercept = opt, color = time), size = .6) +
  theme_ridges(grid = FALSE) +
  scale_fill_brewer(type = "qual") +
  scale_color_brewer(type = "qual") +
  labs(
    title = "Codon optimality regulation is present",
    subtitle = "p-value < 2e-16"
  )

```

Compute a p.value

```{r}
lm(optimality ~ time, data = d) %>% 
  broom::tidy()
```



### Can we see miR-430 effect

```{r}
fit <- glm(mir430mer6 ~ time , data = d, family=binomial(link='logit'))
broom::tidy(fit)
```

Good! We can see miR-430 regulation.

### EDA plots

```{r}
res <- d %>% 
  mutate(
    time = as.factor(time),
    opt = ntile(optimality, n = 10)
  ) %>% 
  group_by(time, opt) %>% 
  summarise(mir_p = mean(mir430mer6)) %>% 
  ungroup()

res %>% 
  ggplot(aes(x = opt, y = mir_p, color = time)) +
  geom_line() +
  geom_smooth()

res %>% 
  pivot_wider(values_from = mir_p, names_from = time) %>% 
  mutate(
    log2 = log2((`8` + .000001) / (`2` + .000001))
  ) %>% 
  ggplot(aes(x = opt, y = log2)) +
  geom_line() +
  geom_smooth()
```
Cool! The results agreee, but we need the model.

### Boostrap resamples

```{r main, message=F, fig.height=4, fig.width=5}
boostrap_sample <- function(i) {
  sample_frac(d, prop = 1, replace = T) %>% 
    mutate(
      time = as.factor(time),
      opt = ntile(optimality, n = 7)
    ) %>% 
    group_by(time, opt) %>% 
    summarise(mir_p = mean(mir430mer6)) %>% 
    ungroup() %>% 
    mutate(boot = i)
}


resultado <- 1:100 %>% 
  map_df(boostrap_sample)

resultado %>% 
  pivot_wider(values_from = mir_p, names_from = time) %>% 
  mutate(
    log2 = log2((`8` + .000001) / (`2` + .000001))
  ) %>% 
  ggplot(aes(x = opt, y = log2)) +
  geom_hline(yintercept = 0) +
  geom_boxplot(aes(group = opt, fill = opt), outlier.shape = NA) +
  scale_x_continuous(breaks = 1:15) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 3.5) +
  labs(
    x = "optimality quantiles",
    y = "log2 fold-change\n depletion miR-430 seed",
    title = "Arielome data",
    subtitle = "boostrap resamples"
  ) +
  ggridges::theme_ridges() +
  theme(legend.position = "none")

```


