---
title: "Compute CSC"
author: "Santiago Medina"
date: "1/17/2019"
output: 
  html_document:
    keep_md: true
---

## CSC analysis

Here, I estimate the CSC as defined in [Prensyak et al 2016](https://www.cell.com/cell/fulltext/S0092-8674(15)00195-6) for the a-amanitin time course.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )
# libraries need it for analysis
library(tidyverse)
library(broom)

theme_set(theme_bw())
```


I filter the data $\alpha > 0$, bassically drop low expressed genes.

```{r load_data}
decay <- read_csv("../19-01-11-GetDecayRateFromTimeCourse/results_data/estimated_decay_rates.csv")

decay_rate <- decay %>% 
  select(Gene_ID:estimate) %>% 
  spread(key = term, value = estimate) %>% 
  filter(alpha > 0)

codon_composition <- read_csv("../../data/19-01-17-Get-ORFS-UTRS-codon-composition/sequence-data/zfish_codon_composition.csv")

# normalize so rows add up to 1

codon_probas <- 
  codon_composition %>% 
  gather(key = codon, value = freq, -id_col) %>% 
  group_by(id_col) %>% 
  mutate(p = freq / sum(freq)) %>% 
  ungroup() %>% 
  select(-freq) %>% 
  spread(key = codon, value = p) %>% 
  rename(Gene_ID = id_col)
```


### Data Exploration

```{r vis}
data <- 
  decay_rate %>% 
  select(-alpha) %>% 
  inner_join(codon_probas)

ggplot(data, aes(beta)) +
  geom_histogram(bins = 100) +
  labs(
    title = "Decay Rate Distribution"
  )
```

There is an unpublished a-amanitin data set from Ariel's previous lab. I will load this data and see if they correlate.


```{r cor_with_old_data}
old_hl <- read_tsv("../../data/19-01-17-EMBO2016DATA/datasets/Half_life_Zebrafish.txt")

old_hl %>% 
  rename(Gene_ID = gene) %>% 
  inner_join(data) %>% 
  ggplot(aes(-log(2) / HL, beta)) +
  geom_point(shape = ".") +
  ggpubr::stat_cor() +
  geom_smooth(method = "lm", size = 1/2) +
  labs(
    x = "Decay Rate (old data set, unpublished)",
    y = "My decay rate (new data)",
    title = "Do we get similar estimates for mRNA stability?"
    )
```

GOOD!, the correlation is not bad.

## Estimate CSC values

```{r csc}
csc <- 
  data %>% 
  gather(key = codon, value = proba, -beta, -Gene_ID) %>% 
  select(-Gene_ID) %>% 
  group_by(codon) %>% 
  nest() %>% 
  mutate(
    cor = map(data, ~cor.test(.$beta, .$proba)),
    tidy_cor = map(cor, ~tidy(.))
  ) %>% 
  unnest(tidy_cor) %>% 
  select(codon, estimate:conf.high)

csc %>% 
  ggplot(aes(x=reorder(codon, -estimate), y=estimate)) +
  geom_point(size = 1/2) +
  geom_hline(yintercept = 0, linetype= 3) +
  geom_linerange(aes(ymin=conf.low, ymax=conf.high)) +
  theme(axis.text.x = element_text(angle = 90, size = 4)) +
  labs(
    y = "CSC",
    x = NULL
  )
```

#### Do these estimates correlate with the EMBO 2016 paper estimates?


```{r csc_cors}
read_tsv("../../data/19-01-17-EMBO2016DATA/datasets/Decay_HF_vs_Arielome.txt") %>% 
  rename(codon = Codon) %>% 
  inner_join(csc) %>% 
  ggplot(aes(InVivo_Zebrafish, estimate)) +
  geom_point() +
  ggpubr::stat_cor() +
  geom_smooth(method = "lm") +
  labs(
    x = "CSC EMBO 2016",
    y = "new data CSC"
  )
```

