---
title: "Generate log2 Fold Change Data"
author: "Santiago Medina"
date: "2/5/2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )

library(tidyverse)
theme_set(theme_light())
```

The goal of this notebook is to generate the log2 Fold Change data, this data will be usefull in other analysis.



```{r load}
tc_alpha <- read_csv("../19-01-09-RNAseqProfilesFish/rna-seq-profiles/alpha-amanitin-prolife.csv") 
tc_polya <- read_csv("../19-01-09-RNAseqProfilesFish/rna-seq-profiles/polyA-profile.csv")
tc_ribo <- read_csv("../19-01-09-RNAseqProfilesFish/rna-seq-profiles/ribo-zero-profile.csv")
```

```{r computelog2FC}
## use a function factory
logFCnormalizer <- function(time_var) {
  function(x) {
    log2(x / time_var)
  }
}

test <- 
  tc_alpha %>% 
  mutate_if(is.numeric, function(x) logFCnormalizer(.$s_alpha_2h)(x))

## test check that output is correct

test_check <- any(
  !test$s_alpha_7.5h == (log2(tc_alpha$s_alpha_7.5h / tc_alpha$s_alpha_2h)),
  na.rm = T
)

if (test_check) stop("you did something wrong")

# cool function works, now apply
fc_alpha <- tc_alpha %>% 
  mutate_if(
  .predicate = is.numeric,
  .funs = function(x) logFCnormalizer(.$s_alpha_2h)(x)
)

fc_polya <- tc_polya %>% 
  mutate_if(
  is.numeric,
  function(x) logFCnormalizer(.$s_wt_2h)(x)
)

fc_ribo <- tc_ribo %>% 
  mutate_if(
  is.numeric,
  function(x) logFCnormalizer(.$s_wt_2h)(x)
)


## tidy data to put it together

tidy_fc <- function(fc_tc, .sample_condition) {
  fc_tc %>% 
    select(-Name) %>% 
    gather(key = tmpvar, value = log2FC, -Gene_ID) %>% 
    mutate(
      time = str_extract(tmpvar, "\\d\\.?\\d?") %>% as.numeric()
    ) %>% 
    select(-tmpvar) %>% 
    mutate(sample_condition = .sample_condition)
}

fc_time_course <- bind_rows(
  tidy_fc(fc_ribo, .sample_condition = "wt_ribo"),
  tidy_fc(fc_polya, .sample_condition = "wt_polya"),
  tidy_fc(fc_alpha, .sample_condition = "aamanitin_polya")
) %>% 
  filter(time >= 2) # lower time points do not make sense

## let's plot 100 random genes

fc_time_course %>% 
  filter(Gene_ID %in% sample(Gene_ID, 1000)) %>% 
  ggplot(aes(x = time, y = log2FC, group = Gene_ID)) +
  geom_line(alpha = 1/10) +
  facet_grid(~sample_condition) +
  labs(title = "log2FC late vs early")
```


I will ad an indicator column if the gene is maternal. I have a list of genes that are supposly maternal (*../19-01-17-EMBO2016DATA/datasets/Half_life_Zebrafish.txt*) although I am not 100% sure.

```{r maternals, fig.width=6, fig.height=2.5}
maternals <- read_tsv("../19-01-17-EMBO2016DATA/datasets/Half_life_Zebrafish.txt") %>% 
  pull(gene)

fc_time_course <- mutate(fc_time_course, is_maternal = Gene_ID %in% maternals)

# repeat the plot above, is there a pattern
fc_time_course %>% 
  ggplot(aes(x = as.character(time), y = log2FC, fill = is_maternal)) +
  geom_boxplot(outlier.shape = NA, alpha = 1/2) +
  ggthemes::scale_fill_colorblind() +
  facet_wrap(. ~ sample_condition, scales = "free") +
  coord_cartesian(ylim = c(-7, 7)) +
  theme(legend.position = "bottom")

```

Save the data, I will also save a tidy version of the time course data in case i want to see effect of regulatory sequences at the RNA level.

```{r savedata}

write_csv(fc_time_course, "data/log2FC_earlyVSlate_tidytimecourse.csv")

# here I will also generate a tidy version of the time course
# in case i want to compare mRNA levels, i can resuse the same function

time_course_tidy <- bind_rows(
  tidy_fc(tc_ribo, .sample_condition = "wt_ribo"),
  tidy_fc(tc_polya, .sample_condition = "wt_polya"),
  tidy_fc(tc_alpha, .sample_condition = "aamanitin_polya")
) %>% 
  rename(TPM = log2FC) %>% # in this case the values are TPMs not log2FC
  mutate(is_maternal = Gene_ID %in% maternals)

write_csv(time_course_tidy, "data/RNAseq_tidytimecourse.csv")
```

```{r info}
sessionInfo()
```





