---
title: "Plot k-mer profiles"
author: "Santiago Medina"
date: "2/5/2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )

library(tidyverse)
theme_set(theme_light())
```

### Overview

In this notebook, I will plot the regulatory profile in the new RNA-seq data of the k-mers that were proviously identified.

```{r load_data}
utrs <- read_csv("../../data/19-01-17-Get-ORFS-UTRS-codon-composition/sequence-data/fish_seq_data_cds_3utr.csv") %>% 
  select(-coding) %>% 
  rename(Gene_ID = ensembl_gene_id)
fc_tc <- read_csv("../../data/19-02-05-FoldChangeData/data/log2FC_earlyVSlate_tidytimecourse.csv")

```


```{r old_candidate_kmers_profilies, fig.height=10, fig.width=9}
x <- 
  utrs %>% 
  mutate(
    MiR430 = str_count(`3utr`, "GCACTT"),
    unstable_CTGCTA = str_count(`3utr`, "CTGCTA"),
    unstable_CGGTGA = str_count(`3utr`, "CGGTGA"),
    unstable_GTGCGT = str_count(`3utr`, "GTGCGT"),
    unstable_GGACCG = str_count(`3utr`, "GGACCG"),
    stable_CGATCG = str_count(`3utr`, "CGATCG"),
    stable_TCCCGT = str_count(`3utr`, "TCCCGT")
  ) %>% 
  select(-`3utr`) %>% 
  inner_join(fc_tc) %>% 
  gather(key = kmer, value = n_sites, -log2FC, -time, -Gene_ID, -sample_condition, -is_maternal)


x %>% 
  filter(!is.infinite(log2FC)) %>% 
  group_by(kmer, time, sample_condition) %>% 
  mutate(log2FC = log2FC - mean(log2FC, na.rm = T)) %>% 
  filter(is_maternal, time %% 1 == 0, time > 3) %>% 
  ggplot(aes(x = as.character(time), y = log2FC, fill = n_sites > 0)) +
  geom_boxplot(outlier.shape = NA, size = 1/4) +
  facet_grid(kmer~sample_condition) +
  coord_cartesian(ylim = c(-6, 5.5)) +
  scale_fill_manual(values = c("#855C75", "#D9AF6B")) +
  labs(
    title = "Candidate Regulatory k-mers Identified in previous analysis",
    subtitle = "only maternal genes are shown",
    x = "time (hrs)",
    y = "log2FC late / early"
    )
```

### Conclusions

We can see that the two stabilizing elements look promising and the un-stabilizing are not. After a discussion with Ariel, I find a new model that can be extremely useful to identify potential regulatory factors. This model has discussed it in the lab notebook.

