---
title: "Visualize Models, Identify Potential k-mers"
author: "Santiago Medina"
date: "2/8/2019"
output: 
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, collapse = T,
  fig.path = "./figures/02_", dev = c("png", "pdf"),
  dpi=300,
  fig.height = 3,
  fig.width = 5
  )
library(tidyverse)
library(gridExtra)
theme_set(theme_light())
```


## Overview

```{r load_results, eval=FALSE}
## remove eval false to collect results
deviances <- 
  list.files("results_data/test_kmers/mdl_deviance", pattern = "waic", full.names = T) %>% 
  map_df(read_csv)

params <- 
  list.files("results_data/test_kmers/mdl_deviance", pattern = "params", full.names = T) %>% 
  map_df(read_csv)

## save the results
write_csv(params, "results_data/estimated_parameters_6mers.csv")
write_csv(deviances, "results_data/waic_6mers.csv")
```

```{r relode}
params <- read_csv("results_data/estimated_parameters_6mers.csv")
deviances <- read_csv("results_data/waic_6mers.csv")
```


## plot parameters of default model

```{r default_model}
dflt <- read_csv("results_data/test_kmers/mdl_deviance/GCACTT_params.csv")
mir_waic <- read_csv("results_data/test_kmers/mdl_deviance/GCACTT_waic.csv") %>% 
  filter(val == "waic")
dflt %>% 
  filter(
    parameter %in% c("b_PLS1", "b_PLS2", "b_GCACTT", "b_time:GCACTT", "b_time:PLS1", "b_time:PLS2")
  ) %>% 
  ggplot(aes(x=Estimate, y=reorder(parameter, Estimate))) +
  geom_point(shape=1) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5, height = 0)) +
  geom_vline(xintercept = 0, linetype = 3) +
  facet_grid(~sample_name)
```


```{r eda}
## deviance distribution
deviances %>% 
  filter(val == "waic") %>% 
  filter(sample_name != "aamanitin_polya") %>% 
  ggplot(aes(Estimate)) +
  geom_histogram(bins = 50) +
  facet_grid(~sample_name, scales = "free_x") +
  labs(
    x = "WAIC",
    title = "model deviances"
  ) + 
  geom_vline(
    data = filter(mir_waic, sample_name != "aamanitin_polya"),
    aes(xintercept = Estimate), linetype = 3
    ) +
  scale_x_log10()
```



```{r mdlsummary1}

# extract parameter associated with the k-mer
b_mer <- 
  params %>% 
  filter(
    str_detect(parameter, "b_[ACGT]+"),
    str_extract(parameter, "[ACGT]+") == kmer,
  ) %>% 
  rename_if(is.numeric, .funs = function(x) str_c("b_mer_", x))

# time interaction parameter
b_mer_time <- 
  params %>% 
  filter(
    str_detect(parameter, "b_time:[ACGT]+"),
    str_extract(parameter, "[ACGT]+") == kmer,
  ) %>% 
  rename_if(is.numeric, .funs = function(x) str_c("b_time_", x))

# extract the waic
waics <- deviances %>% 
  filter(val == "waic") %>% 
  rename(waic =Estimate, waic_se = SE) %>% 
  select(-val)


mdl_res <- 
  inner_join(b_mer, b_mer_time, by = c("sample_name", "kmer")) %>% 
  inner_join(waics, by = c("sample_name", "kmer"))
  
mdl_res %>% 
  ## standarize waic just for visualization porposes
  group_by(sample_name) %>% 
  mutate(waic = as.numeric(scale(waic))) %>% 
  filter(sample_name != "aamanitin_polya") %>% 
  ggplot(aes(b_time_Estimate, b_mer_Estimate, color = waic)) +
  geom_point(size = 1/2) + 
  geom_point(
    data = filter(mdl_res, kmer == "GCACTT", sample_name != "aamanitin_polya"),
    color = "black",
    size = 3,
    shape = 1
  ) +
  scale_color_viridis_c(option = "A") +
  geom_vline(xintercept = 0, linetype = 3) +
  geom_hline(yintercept = 0, linetype = 3) +
  facet_wrap(~sample_name, scales = "free_x") +
  labs(
    y = "k-mer affect at 6 hrs",
    x = "k-mer/time interaction effect",
    title = "6-mers profiles",
    color = "WAIC\n(scaled)"
  )
```

NOTE: In general there seems to be a positive effect of each k-mer. This maybe because of the UTR length affects mRNA stability as suggested by Mishima 2016. i may need to add a second model where I include the UTR length.


correlate wild type effect with ribo0 effect.

```{r corre_wt_polya}
mdl_res %>% 
  select(b_mer_Estimate, sample_name, kmer) %>% 
  filter(sample_name != "aamanitin_polya") %>% 
  spread(key = sample_name, value = b_mer_Estimate) %>% 
  ggplot(aes(wt_polya, wt_ribo, label = kmer)) +
  geom_point(size = 1/4) +
  geom_abline(linetype = 3, color = "steelblue")
```


## Candidates

What is a candidata k-mer?

+ effect in ribo 0
+ effect in poly A
+ WAIC better than the null model

By effect I mean that the 95% of the posterior estimate does not contain 0.

I will write a function to test each element.

```{r test_kmer} 
kprams <- 
  params %>% 
  filter(
    str_detect(parameter, "b_[ACGT]{6}") |  str_detect(parameter, "b_time:[ACGT]{6}"),
    str_extract(parameter, "[ACGT]{6}") == kmer
  )


test_kmer <- function(kmer_data) {
 # returns k-mer state
  is_positive <- 
    kmer_data %>% 
    filter(sample_name != "aamanitin_polya") %>% 
    select( Q2.5:Q97.5) %>% 
    gather() %>% 
    pull(value) %>% 
    (function(x) (x>0))() %>% 
    all()
  
  down_polya <- kmer_data %>% 
    filter(sample_name == "wt_polya") %>% 
    select(Q2.5, Q97.5) %>% 
    gather() %>% 
    pull(value) %>%
    (function(x) (x<0))() %>% 
    all()
  
  down_ribo <- kmer_data %>% 
    filter(sample_name == "wt_ribo") %>% 
    select(Q2.5, Q97.5) %>% 
    gather() %>% 
    pull(value) %>%
    (function(x) (x<0))() %>% 
    all()
  
  if(!any(c(is_positive, down_polya, down_ribo))) return("no effect")
  if (is_positive) return("stabilizing")
  return("destabalizing")
  
}

candidates <- 
  kprams %>% 
  group_by(kmer) %>% 
  nest() %>% 
  mutate(effect = map_chr(data, test_kmer)) %>% 
  select(-data)

## there seems to be some sequence that are not very significant
## at the poly-A but show strong effect at 6
other_candidates <- filter(b_mer, b_mer_Estimate > 0.075, sample_name == "wt_polya")$kmer

write_csv(candidates, "results_data/candidates_6mers.csv")
cat("How many candidates?", sum(candidates$effect != "no effect"))
```

```{r showcandidates}
mdl_res %>% 
  ## standarize waic just for visualization porposes
  filter(sample_name != "aamanitin_polya") %>% 
  inner_join(candidates) %>% 
  filter(kmer != "GCACTT") %>% 
  ggplot(aes(b_time_Estimate, b_mer_Estimate, color = effect, size = effect)) +
  geom_point() + 
  geom_vline(xintercept = 0, linetype = 3) +
  geom_hline(yintercept = 0, linetype = 3) +
  facet_wrap(~sample_name, scales = "free_x") +
  scale_size_manual(values = c(1, .005 ,1)) +
  scale_color_manual(values = c("red", "grey", "blue")) +
  labs(
    y = "k-mer affect at 6 hrs",
    x = "k-mer/time interaction effect",
    title = "6-mers profiles",
    color = "candidate k-mer"
  ) +
  theme(legend.position = "none")
```

## Candidate sequences

Plot profiles effects

```{r candidates_heatmpa, fig.height=6.5, fig.width=3}
timeh <- tibble(t = 4:7, t_scaled =  as.numeric(scale(t)))

effect_equation <- function(b_kmer, b_kmer_time) {
  function(t) {
    b_kmer + b_kmer_time * t
  }
}

get_effects_kmer <- function(.kmer) {
  filter(params, kmer == .kmer) %>% 
    filter(
      str_extract(parameter, "[ACGT]+") == kmer,
      kmer == .kmer
    ) %>% 
    select(
      parameter, Estimate, sample_name
    ) %>%
    mutate(parameter = map_chr(parameter, ~if_else(str_detect(., "time"), "interaction", "effect_b"))) %>% 
    spread(key = parameter, value = Estimate) %>% 
    mutate(p_eq = map2(effect_b, interaction, effect_equation)) %>% 
    mutate(effect = map(
      p_eq,
      function(x) mutate(timeh, effect_kmer = map_dbl(t_scaled, x))
    )) %>% 
    select(sample_name, effect) %>% 
    unnest(effect)
  
}


dp <- waics %>% 
  group_by(kmer) %>% 
  mutate(waic = as.numeric(scale(waic))) %>% 
  summarise(mean_waic = mean(waic)) %>% 
  inner_join(candidates) %>% 
  filter(effect != "no effect") %>% 
  mutate(ef = map(kmer, get_effects_kmer)) %>% 
  unnest(ef)

order_mers <- dp %>% 
  filter(t == 7, sample_name == "wt_ribo") %>% 
  arrange(-effect_kmer) %>% 
  pull(kmer)

dp %>% 
  mutate(kmer = factor(kmer, levels = order_mers)) %>% 
  ggplot(aes(y=kmer, x=t, fill = effect_kmer)) +
  geom_tile() +
  facet_grid(effect~sample_name, space="free", scales = "free") +
  scale_fill_gradient2(low = "red", high = "steelblue", mid = "white", midpoint = 0) +
  scale_x_continuous(expand = c(0, 0), breaks = 4:7) +
  scale_y_discrete(expand = c(0, 0)) +
  labs(
    x = "time (hrs)",
    title = "candidate regulatory 6-mers",
    color = "effect"
  ) +
  theme(axis.text.y = element_text(size = 4),
        axis.text.x = element_text(size = 4),
        strip.text.x = element_text(size = 4))
```

